Global patterns of forest autotrophic carbon fluxes

### 2020-08-03 manuscript/manuscript draft.Rmd
merging versions

\newpage
###Abstract
Carbon (C) fixation, allocation, and metabolism by trees set the basis for energy and material flows in forest ecosystems and define their interactions with Earth's changing climate. 
However, we lack a cohesive synthesis on how forest carbon fluxes vary globally with respect to climate and one another. 
Here, we draw upon **#** records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant autotrophic carbon fluxes, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale. 
We show that, across all flux variables analyzed, C cycling decreases linearly with absolute latitude -- a finding that confirms multiple previous studies but contradicts the idea that net primary productivity ($NPP$) of temperate forests rivals that of tropical forests. 
C flux variables generally displayed similar trends across latitude and multiple climate variables, with few differences in allocation detectable at this global scale.
**Climate explained a significant proportion (#-#%) of variation in all C fluxes analyzed**, with temperature variables in general and mean annual temperature (MAT) in particular being the best predictors of C flux on this global scale. 
The effects of temperature were modified by moisture availability, with C flux reduced under hot and dry conditions and sometimes under very high precipitation. 
FACF declined with temperature seasonality, but growing season length did not improve upon MAT as a predictor.  
Within the growing season, the influence of climate on C cycling was small but significant for a number of flux variables. 
These findings clarify how forest C flux varies with latitude and climate on a global scale. 
In a period of accelerating climatic change, understanding of the fundamental climatic controls on forest C cycling sets a foundation for understanding patterns of change.

@@

Carbon cycling in forests worldwide provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations [@millennium_ecosystem_assessment_ecosystems_2005], while strongly influencing atmospheric CO~2~ and climate [@bonan_forests_2008]. Forests' autotrophic carbon fluxes (FACF)--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance [@zak_plant_1994; @niedzialkowska_species_2010] and possibly diversity [@waide_relationship_1999; @chu_direct_2018]. FACF influences all organic matter stocks in forest ecosystems and is linked to cycling of energy, water, and nutrients.

Critically, FACF also define forest interactions with Earth's changing climate. Over 69 Gt of CO~2~ cycle through Earth's forests each year [@badgley_terrestrial_2019], and in recent decades their net C sequestration (~2.4 Gt C yr^-1^) offset roughly 30% of anthropogenic fossil fuel emissions [@pan_large_2011]. As atmospheric carbon dioxide levels continue to rise, driving climate change, forests will play a critical role in shaping the future of Earth's climate [@cavaleri_urgent_2015; @rogelj_mitigation_2018]. However, our ability to draw general macroscopic conclusions regarding global variation in FACF with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables; and typically consider only one or a few variables at a time.

FACF decrease with latitude, but it remains unclear whether and how the shape of this relationship varies between fluxes. FACF are lowest in the boreal regions, and increase into the temperate regions [@luyssaert_co_2007; @huston_global_2009; @beer_terrestrial_2010; @piao_forest_2010; @jung_global_2011]. However, evidence is inconclusive on whether primary productivity continues to increase into the tropics. Evidence for this is further complicated by the fact that different studies use different measures of productivity to explore these relationships. For example, modelling of global terrestrial ecosystem gross primary productivity ($GPP$) through upscaling and calibration of eddy flux measurements indicates that GPP peaks in tropical forests [@beer_terrestrial_2010; @jung_global_2011; @badgley_terrestrial_2019; @li_mapping_2019]. In contrast, some studies suggest that the highest values of net primary productivity ($NPP$) may be found in temperate forests [@luyssaert_co_2007; @huston_global_2009], while others find $NPP$ highest in the tropics and decreasing with latitude [@simova_enigma_2017]. Other studies have found that aboveground net primary productivity ($ANPP$) decreases weakly with latitude [@huston_global_2009; @gillman_latitude_2015].

The latitudinal gradient in FACF, along with altitudinal gradients [@girardin_net_2010; @malhi_variation_2017], is primarily driven by climate, which is a significant driver of FACF across broad spatial scales [@luyssaert_co_2007; @cleveland_relationships_2011; @hursh_sensitivity_2017]. The majority of studies have focused on exploring the relationships of FACF to mean annual temperature (MAT) and mean annual precipitation (MAP), as the most commonly reported site-level climate variables. There is strong evidence that both MAT and MAP show significant positive relationships with FACF [@chu_does_2016]. However, as with latitude, the shape of those relationships is not always clear, and, again, is complicated by the use of different measures of FACF across studies. Various measures of primary productivity {**FACF?**} saturate at high levels of MAP, though the saturation points identified vary from 1500mm [@luyssaert_co_2007] up to 2445mm MAP [@schuur_productivity_2003]. Studies of the influence of MAT on productivity {**FACF?**} are less conclusive. Luyssaert et al. [-@luyssaert_co_2007] examined GPP and NPP and found that, while $GPP$ increases linearly with MAT, NPP saturates at around 10$^\circ$C MAT. In contrast, Larjavaara and Muller-Landau [-@larjavaara_temperature_2012], find that increases in $GPP$ saturate at approximately 25$^\circ$C MAT, while Schuur [-@schuur_productivity_2003] finds that $NPP$ increases linearly with temperature. Taylor et al. [-@taylor_temperature_2017] showed a positive interaction between MAT and MAP in shaping tropical forest productivity, such that high rainfall had a negative effect on productivity in cooler climates, compared to a positive effect in warmer climates. Such complicated dynamics play out with belowground fluxes as well, with typically dominant soil temperature mediated by moisture and carbon supply [@hursh_sensitivity_2017; @xu_contribution_2016].

MAT and MAP are very coarse measures of climate, and so fail to capture much variation in climate on an intra-annual scale, including the effects of factors such as growing season length, number of frost-free days, temperature seasonality, and dry season length. Some studies have suggested that the apparently strong relationship between MAT and FACFs is actually a consequence of the correlation between MAT and growing season length [@kerkhoff_plant_2005; @michaletz_convergence_2014; @michaletz_drivers_2018]. Kerkhoff et al. [-@kerkhoff_plant_2005] and Michaletz et al. [-@michaletz_convergence_2014] find that, within the growing season, there is no significant relationship between net primary productivity (loosely defined) and MAT, suggesting that the influence of temperature may be limited to determining the length of the frost-free growing season.

**the following doesn't fit very well...**  
In addition, FACF can be influenced by other, often interactive, factors across a range of scales [@cleveland_relationships_2011]. On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], biodiversity [@liang_positive_2016], management [@simova_enigma_2017], and nutrient availability [@aragao_above_2009]. There is evidence that FACFs also respond to variables such as cloud cover [@taylor_temperature_2017], solar radiation [@fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005].

The recent development of the Global Forest Carbon database (ForC), which synthesizes multiple variables and including records of stand history [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018], opens up the possibility for a standardized analysis of global scale variation in multiple FACF and the principle climatic drivers of these patterns. In order to approach this broad topic, we simplify the major gaps in our knowledge to five broad questions and corresponding hypotheses (Table 1). First, we ask how nine FACF in ForC vary with latitude. We then test how these fluxes relate to MAT and MAP, and additionally how they respond to other, less well-studied, climate variables. Finally, we consider the relationship between FACF and seasonality, considering the role of seasonality in explaining variation in carbon fluxes, and the influence of climate on FACF standardized by growing season length.

@@

This analysis focused on nine FACF included in the open-access ForC database (Table 2) [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018]. ForC contains records of field-based measurements of forest carbon stocks and annual fluxes, compiled from original publications and existing data compilations and databases. Associated data, such as stand age, measurement methodologies, and disturbance history, are also included. The database was significantly expanded since the publication of @anderson-teixeira_forc_2018 through integration with the Global Soil Respiration Database [@bond-lamberty_global_2010]. Additional targeted literature searches were conducted to identify any further available data on the FACF analyzed here, with particular focus on mature forests in temperate and boreal regions, which were not included in the review of @anderson-teixeira_carbon_2016. We used ForC v3.0, archived on Zenodo with DOI 10.5281/zenodo.3403855. This version contained 29,730 records from 4,979 plots, representing 20 distinct ecozones across all forested biogeographic and climate zones.

This analysis focused on mature forests with no known history of signficant disturbance or management. 
There is evidence that stand age influences patterns of FACF and carbon allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@de_lucia_forest_2007; @gillman_latitude_2015]. To reduce any biasing effects of stand age, we included only stands of known age $\ge$ 100 years and those described by terms such as "mature", "intact", or "old-growth". 
Since management can alter observed patterns of FACF [@simova_enigma_2017], sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or including the term "managed" in their site description. 
Sites that had experienced significant disturbance within the past 100 years were also excluded. Disturbances that qualified sites for exclusion included major cutting or harvesting, burning, flooding, drought and storm events with site mortality >10% of trees. Grazed sites were retained.

```{r eval = TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
FACF_table <- read.csv("FACF_variables.csv", stringsAsFactors = FALSE, check.names = FALSE)
names(FACF_table) [6]  <- paste0(names(FACF_table) [6], footnote_marker_symbol(1))
kable(FACF_table, format = "latex", booktabs = TRUE, caption = "Definitions and sample sizes of FACF variables used in analysis. All variables are in units of Mg C ha$^{-1}$ yr$^{-1}$. ", escape = F) %>%
  kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
  column_spec(2, width = "4cm") %>%
  column_spec(3, width = "7cm") %>%
  column_spec(4, width = "10cm") %>%
  column_spec(5:6, width = "1.5cm") %>%
  add_header_above(c(" ",  " "," ",  " ", "Sample size" = 2)) %>%
  kableExtra::footnote(symbol = c("Geographic areas group geographically proximate sites, defined using a hierarchical cluster analysis on the distance matrix of the sites, and a cutoff of 25km"), footnote_as_chunk = FALSE) 
```

@@

The effects of latitude and climate on FACF were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r:_2018]. The basic model for all analyses included a fixed effect of latitude or climate and a random effect of plot nested within geographic area. Geographic areas--*i.e.*, spatially clustered sites--are defined within ForC using a hierarchical cluster analysis on the distance matrix of the sites and a cutoff of 25km [@anderson-teixeira_forc_2018]. We experimented with inclusion of altitude as a fixed effect, but excluded it from the final models because it added very little explanatory power--that is, the difference in AIC ($\Delta AIC$) relative to models excluding altitude was generally small (often $\Delta AIC$<2). Effects were considered signficant when inclusion of the term of interest resulted in $\Delta AIC \ge 2.0$ relative to a corresponding null model. All $R^2$ values presented here are marginal $R^2$ values, and refer to the proportion of variation explained by only the fixed effects. Specific analyses are as described below.

We first examined the relationship between latitude and FACF (Q1; Table 1). We tested models with latitude as a first-order linear term (corresponding null: model without latitude) and as a second-order polynomial or logarithmic term (corresponding null: model with latitude as a first-order linear term). We selected as the best model that with the highest $\Delta AIC$ relative to a null model with no fixed term.

@@

We next examined the relationships of FACF to climate variables (Q2-Q4; Table 1). As with latitude, we tested both linear and polynomial fits for each climate variable. We tested relationships of each FACF (Table 2) against each climate variable (Table S1), but focus presentation--including focal hypotheses (Table 1)--on climate variables that explained >20% of variation in FACF. **Becky, please make this more specific. I drafted this based on the results, but its not very clear.** **BBL: I agree, not clear and might trigger reviewer concerns about tailoring hypotheses to significance.**

@@

To test whether and how FACF varied with climate when standardised by growing season length (Q5), we first standardized FACF by dividing by growing season length (as defined above). We then tested for correlations between these standardised fluxes and growing season climate variables, using only first-order linear models.

@@

In total, we analyzed 1228 records from nine FACF variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years. These records represented a total of **#** plots in 154 distinct geographic areas across all forested biogeographic and climate zones (Fig. 1, Table 2).

*How do FACF vary with latitude?*

All major carbon fluxes decreased linearly with latitude (Fig. 2; Table S2). Latitude was a strong predictor for many of the carbon fluxes--particularly the larger fluxes (Table S2). Specifically, latitude explained 64% of variation in GPP (n = 254, p<0.0001), 50% in NPP (n = 114, p<0.0001) and 45% in ANPP (n = 259, p<0.0001). The FACF that were most poorly predicted by latitude were $BNPP_{fine.root}$ ($R^2$=0.17) and $ANPP_{stem}$ ($R^2$=0.18). For all FACF, the relationship with latitude was better fit by the first-order linear model than by second-order polynomial or logarithmic models.

@@

There was little evidence of systematic variation in C allocation with latitude or climate (Fig. S3). Of 28 relationships tested (7 FACF ratios regressed against latitude, MAT, MAP, and temperature seasonality), there were only five signficant relationships, all with $R^2 \le 0.4$ (Fig. S3). The proportion of NPP allocated to $ANPP_{foliage}$ decreased with latitude ($R^2=0.32$) and increased with MAT ($R^2=0.37$). The proportion of $NPP$ allocated aboveground ($ANPP$) decreased weakly with latitude ($R^2=0.11$) and temperature seasonality ($R^2=0.17$), while increasing with MAT ($R^2=0.11$).

*How do FACF relate to MAT and MAP?*

@@

MAP was a significant (p<0.05) predictor of all fluxes (Figs. 4a, S4-S5; Table S2). 
However, it explained little variation: with the exception of $R_{auto}$, MAP explained at most 25% of variation in FACF. 
For five of the nine fluxes, a second-order polynomial model provided the best fit, while first-order linear and lognormal relationships provided each provided the best fit for two flxues. 
All fluxes increased with MAP up to at least 2000 mm, above which responses were variable (Figs. 4, S4-S5).

@@

```{r echo=FALSE, out.width='100%', fig.cap = "Interactive effects of mean annual temperature and mean annual precipitation on FACF. For visualization purposes, data points are grouped into bins of 0 - 1000, 1001 - 2000, 2001 - 3000, and >3000mm mean annual precipitation, and lines of best fit models are plotted for mean annual precipitation values of 500, 1500, 2500, and 3500mm. All regressions are significant $(p<0.05)$.", fig.pos='H'}
  knitr::include_graphics("mat_map_interaction.png")
```

*How do FACF relate to other climate variables?*

Our results indicated that FACF were most strongly explained by temperature at the global scale, with temperature-related climate variables coming out as strong predictors of FACF. In addition to MAT, several of its correlates (Fig. S2) were consistently identified as strong univariate predictors of FACF: temperature seasonality, annual temperature range, annual frost days, PET, and length of growing season (Figs. 4, S4-S7).

We found a significant relationship between C flux and potential evapotranspiration for all FACF. $ANPP_{foliage}$, $BNPP_{fine.root}$ and $R_{root}$ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (Fig. 4c, S4-5; Table S2). We found strong evidence for a saturation point or peak with PET: FACF tended to increase at values below 1000mm, before saturating between 1200 and 1700mm. There was also evidence that some FACF begin to decrease at values above 1800mm PET.

Vapour pressure deficit was a significant predictor of C flux for all FACF. $BNPP_{fine.root}$ showed a linear relationship with vapour pressure deficit ($R^2$ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (Figs. 4d, S4-5; Table S2). FACF initially increased with vapour pressure deficit, before saturating at around 0.8 kPa, after wich point they began to decrease.

All fluxes, with the exception of $R_{root}$, showed a positive linear relationship with solar radiation (Figs. S4-S5, Table S2). Solar radiation explained a low proportion of variability in all FACF, explaining less than 20% of the variation in each flux, with the exception of $R_{auto}$ ($R^2$ = 0.26, p<0.05).

Annual wet days, cloud cover, and aridity were poor or non-significant explainers of variation in FACF, explaining less than 20% of the variation in each of the carbon fluxes (Figs. S4-S5; Table S2).

@@

*What is the role of seasonality in explaining FACF?*

Temperature seasonality was, overall, the best predictor of FACF. **(update this paragraph based on new results)**
$GPP$, $NPP$, $ANPP$, and $R_{root}$ exhibited a polynomial relationship with seasonality (all p<0.05; Figs. 4e, S6-7; Table S2). $ANPP_{foliage}$, $ANPP_{stem}$ and $R_{auto}$ decreased linearly with temperature seasonality (all p<0.05; Figs. 4e, S6-S7; Table S2). Temperature seasonality was strongly correlated with annual temperature range, which was likewise a similarly strong predictor of FACF (Table S2). 
FACF were highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower. **BBL: perhaps put this into an ecosystem context; what are these? Aseasonal subtropical places?**

In contrast, there was no significant effect of precipitation seasonality on FACF, and both maximum vapour pressure deficit, and water stress months were poor or non-significant predictors of variation in FACF (Figs. S6-S7; Table S2).

>>>>

\newpage
###Abstract
Carbon (C) fixation, allocation, and metabolism by trees set the basis for energy and material flows in forest ecosystems and define their interactions with Earth's changing climate. 
However, we lack a cohesive synthesis on how forest carbon fluxes vary globally with respect to climate and one another. 
Here, we draw upon **#** records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant autotrophic carbon fluxes, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale. 
We show that, across all flux variables analyzed, C cycling decreases linearly with absolute latitude -- a finding that confirms multiple previous studies but contradicts the idea that net primary productivity ($NPP$) of temperate forests rivals that of tropical forests. 
C flux variables generally displayed similar trends across latitude and multiple climate variables, with few differences in allocation detectable at this global scale.
**Climate explained a significant proportion (#-#%) of variation in all C fluxes analyzed**, with temperature variables in general and mean annual temperature (MAT) in particular being the best predictors of C flux on this global scale. 
The effects of temperature were modified by moisture availability, with C flux reduced under hot and dry conditions and sometimes under very high precipitation. 
C fluxes declined with temperature seasonality, but growing season length did not improve upon MAT as a predictor.  
Within the growing season, the influence of climate on C cycling was small but significant for a number of flux variables. 
These findings clarify how forest C flux varies with latitude and climate on a global scale. 
In a period of accelerating climatic change, understanding of the fundamental climatic controls on forest C cycling sets a foundation for understanding patterns of change.

@@

Carbon cycling in forests worldwide provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations [@millennium_ecosystem_assessment_ecosystems_2005], while strongly influencing atmospheric CO~2~ and climate [@bonan_forests_2008]. Forests' autotrophic carbon fluxes--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance [@zak_plant_1994; @niedzialkowska_species_2010] and possibly diversity [@waide_relationship_1999; @chu_direct_2018]. They influence all organic matter stocks in forest ecosystems and is linked to cycling of energy, water, and nutrients.

Critically, autotrophic C fluxes also define forest interactions with Earth's changing climate. Over 69 Gt of CO~2~ cycle through Earth's forests each year [@badgley_terrestrial_2019], and in recent decades their net C sequestration (~2.4 Gt C yr^-1^) offset roughly 30% of anthropogenic fossil fuel emissions [@pan_large_2011]. As atmospheric carbon dioxide levels continue to rise, driving climate change, forests will play a critical role in shaping the future of Earth's climate [@cavaleri_urgent_2015; @rogelj_mitigation_2018]. However, our ability to draw general macroscopic conclusions regarding global variation in C cycling with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables; and typically consider only one or a few variables at a time.

Forest C fluxes decrease with latitude, but it remains unclear whether and how the shape of this relationship varies between fluxes. Fluxes are lowest in the boreal regions, and increase into the temperate regions [@luyssaert_co_2007; @huston_global_2009; @beer_terrestrial_2010; @piao_forest_2010; @jung_global_2011]. However, evidence is inconclusive on whether primary productivity continues to increase into the tropics. Evidence for this is further complicated by the fact that different studies use different measures of productivity to explore these relationships. For example, modelling of global terrestrial ecosystem gross primary productivity ($GPP$) through upscaling and calibration of eddy flux measurements indicates that GPP peaks in tropical forests [@beer_terrestrial_2010; @jung_global_2011; @badgley_terrestrial_2019; @li_mapping_2019]. In contrast, some studies suggest that the highest values of net primary productivity ($NPP$) may be found in temperate forests [@luyssaert_co_2007; @huston_global_2009], while others find $NPP$ highest in the tropics and decreasing with latitude [@simova_enigma_2017]. Other studies have found that aboveground net primary productivity ($ANPP$) decreases weakly with latitude [@huston_global_2009; @gillman_latitude_2015].

The latitudinal gradient in C flux rates, along with altitudinal gradients [@girardin_net_2010; @malhi_variation_2017], is driven primarily by climate, which is a significant driver of C fluxes across broad spatial scales [@luyssaert_co_2007; @cleveland_relationships_2011; @hursh_sensitivity_2017]. The majority of studies have focused on exploring the relationships of C fluxes to mean annual temperature (MAT) and mean annual precipitation (MAP), as the most commonly reported site-level climate variables. There is strong evidence that both MAT and MAP show significant positive relationships with C fluxes [@chu_does_2016]. However, as with latitude, the shape of those relationships is not always clear, and, again, is complicated by the use of different C flux variables across studies. Various measures of primary productivity saturate at high levels of MAP, though the saturation points identified vary from 1500mm [@luyssaert_co_2007] up to 2445mm MAP [@schuur_productivity_2003]. Studies of the influence of MAT on productivity are less conclusive. Luyssaert et al. [-@luyssaert_co_2007] examined GPP and NPP and found that, while $GPP$ increases linearly with MAT, NPP saturates at around 10$^\circ$C MAT. In contrast, Larjavaara and Muller-Landau [-@larjavaara_temperature_2012], find that increases in $GPP$ saturate at approximately 25$^\circ$C MAT, while Schuur [-@schuur_productivity_2003] finds that $NPP$ increases linearly with temperature. Taylor et al. [-@taylor_temperature_2017] showed a positive interaction between MAT and MAP in shaping tropical forest productivity, such that high rainfall had a negative effect on productivity in cooler climates, compared to a positive effect in warmer climates. Such complicated dynamics play out with belowground fluxes as well, with typically dominant soil temperature mediated by moisture and carbon supply [@hursh_sensitivity_2017; @xu_contribution_2016].

MAT and MAP are very coarse measures of climate, and so fail to capture much variation in climate on an intra-annual scale, including the effects of factors such as growing season length, number of frost-free days, temperature seasonality, and dry season length. Some studies have suggested that the apparently strong relationship between MAT and C fluxes is actually a consequence of the correlation between MAT and growing season length [@kerkhoff_plant_2005; @michaletz_convergence_2014; @michaletz_drivers_2018]. Kerkhoff et al. [-@kerkhoff_plant_2005] and Michaletz et al. [-@michaletz_convergence_2014] find that, within the growing season, there is no significant relationship between net primary productivity (loosely defined) and MAT, suggesting that the influence of temperature may be limited to determining the length of the frost-free growing season.

**the following doesn't fit very well...**  
In addition, forest C fluxes can be influenced by other, often interactive, factors across a range of scales [@cleveland_relationships_2011]. On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], biodiversity [@liang_positive_2016], management [@simova_enigma_2017], and nutrient availability [@aragao_above_2009]. There is evidence that C fluxes also respond to variables such as cloud cover [@taylor_temperature_2017], solar radiation [@fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005].

The recent development of the Global Forest Carbon database (ForC), which synthesizes multiple variables and including records of stand history [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018], opens up the possibility for a standardized analysis of global scale variation in multiple C fluxes and the principle climatic drivers of these patterns. In order to approach this broad topic, we simplify the major gaps in our knowledge to five broad questions and corresponding hypotheses (Table 1). First, we ask how nine forest autotropic carbon fluxes in ForC vary with latitude. We then test how these fluxes relate to MAT and MAP, and additionally how they respond to other, less well-studied, climate variables. Finally, we consider the relationship between C flux and seasonality, considering the role of seasonality in explaining variation in carbon fluxes, and the influence of climate on C flux standardized by growing season length.

@@

This analysis focused on nine C flux variables included in the open-access ForC database (Table 2) [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018]. ForC contains records of field-based measurements of forest carbon stocks and annual fluxes, compiled from original publications and existing data compilations and databases. Associated data, such as stand age, measurement methodologies, and disturbance history, are also included. The database was significantly expanded since the publication of @anderson-teixeira_forc_2018 through integration with the Global Soil Respiration Database [@bond-lamberty_global_2010]. Additional targeted literature searches were conducted to identify further available data on the fluxes analyzed here, with particular focus on mature forests in temperate and boreal regions, which were not included in the review of @anderson-teixeira_carbon_2016. We used ForC v3.0, archived on Zenodo with DOI 10.5281/zenodo.3403855. This version contained 29,730 records from 4,979 plots, representing 20 distinct ecozones across all forested biogeographic and climate zones.

This analysis focused on mature forests with no known history of signficant disturbance or management. 
There is evidence that stand age influences patterns of C flux and allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@de_lucia_forest_2007; @gillman_latitude_2015]. To reduce any biasing effects of stand age, we included only stands of known age $\ge$ 100 years and those described by terms such as "mature", "intact", or "old-growth". 
Since management can alter observed patterns of C cycling [@simova_enigma_2017], sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or including the term "managed" in their site description. 
Sites that had experienced significant disturbance within the past 100 years were also excluded. Disturbances that qualified sites for exclusion included major cutting or harvesting, burning, flooding, drought and storm events with site mortality >10% of trees. Grazed sites were retained.

```{r eval = TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
FACF_table <- read.csv("FACF_variables.csv", stringsAsFactors = FALSE, check.names = FALSE)
names(FACF_table) [6]  <- paste0(names(FACF_table) [6], footnote_marker_symbol(1))
kable(FACF_table, format = "latex", booktabs = TRUE, caption = "Definitions and sample sizes of carbon flux variables used in analysis. All variables are in units of Mg C ha$^{-1}$ yr$^{-1}$. ", escape = F) %>%
  kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
  column_spec(2, width = "4cm") %>%
  column_spec(3, width = "7cm") %>%
  column_spec(4, width = "10cm") %>%
  column_spec(5:6, width = "1.5cm") %>%
  add_header_above(c(" ",  " "," ",  " ", "Sample size" = 2)) %>%
  kableExtra::footnote(symbol = c("Geographic areas group geographically proximate sites, defined using a hierarchical cluster analysis on the distance matrix of the sites, and a cutoff of 25km"), footnote_as_chunk = FALSE) 
```

@@

The effects of latitude and climate on C fluxes were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r:_2018]. The basic model for all analyses included a fixed effect of latitude or climate and a random effect of plot nested within geographic area. Geographic areas--*i.e.*, spatially clustered sites--are defined within ForC using a hierarchical cluster analysis on the distance matrix of the sites and a cutoff of 25km [@anderson-teixeira_forc_2018]. We experimented with inclusion of altitude as a fixed effect, but excluded it from the final models because it added very little explanatory power--that is, the difference in AIC ($\Delta AIC$) relative to models excluding altitude was generally small (often $\Delta AIC$<2). Effects were considered signficant when inclusion of the term of interest resulted in $\Delta AIC \ge 2.0$ relative to a corresponding null model. All $R^2$ values presented here are marginal $R^2$ values, and refer to the proportion of variation explained by only the fixed effects. Specific analyses are as described below.

We first examined the relationship between latitude and C fluxes (Q1; Table 1). We tested models with latitude as a first-order linear term (corresponding null: model without latitude) and as a second-order polynomial or logarithmic term (corresponding null: model with latitude as a first-order linear term). We selected as the best model that with the highest $\Delta AIC$ relative to a null model with no fixed term.

@@

We next examined the relationships of C fluxes to climate variables (Q2-Q4; Table 1). As with latitude, we tested both linear and polynomial fits for each climate variable. We tested relationships of each C flux (Table 2) against each climate variable (Table S1), but focus presentation--including focal hypotheses (Table 1)--on climate variables that explained >20% of variation in C fluxes. **Becky, please make this more specific. I drafted this based on the results, but its not very clear.** **BBL: I agree, not clear and might trigger reviewer concerns about tailoring hypotheses to significance.**

@@

To test whether and how C flux varied with climate when standardised by growing season length (Q5), we first standardized annual C flux by dividing by growing season length (as defined above). We then tested for correlations between these standardised fluxes and growing season climate variables, using only first-order linear models.

@@

In total, we analyzed 1228 records from nine forest autotrophic C flux variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years. These records represented a total of **#** plots in 154 distinct geographic areas across all forested biogeographic and climate zones (Fig. 1, Table 2).

*How does C flux vary with latitude?*

All major carbon fluxes decreased linearly with latitude (Fig. 2; Table S2). Latitude was a strong predictor for many of the carbon fluxes--particularly the larger fluxes (Table S2). Specifically, latitude explained 64% of variation in GPP (n = 254, p<0.0001), 50% in NPP (n = 114, p<0.0001) and 45% in ANPP (n = 259, p<0.0001). The C fluxes that were most poorly predicted by latitude were $BNPP_{fine.root}$ ($R^2$=0.17) and $ANPP_{stem}$ ($R^2$=0.18). For all C fluxes, the relationship with latitude was better fit by the first-order linear model than by second-order polynomial or logarithmic models.

@@

There was little evidence of systematic variation in C allocation with latitude or climate (Fig. S3). Of 28 relationships tested (7 ratios among C flux variables regressed against latitude, MAT, MAP, and temperature seasonality), there were only five significant relationships, all with $R^2 \le 0.4$ (Fig. S3). **The proportion of NPP allocated to $ANPP_{foliage}$ decreased with latitude ($R^2=0.32$) and increased with MAT ($R^2=0.37$). The proportion of $NPP$ allocated aboveground ($ANPP$) decreased weakly with latitude ($R^2=0.11$) and temperature seasonality ($R^2=0.17$), while increasing with MAT ($R^2=0.11$).   **

*How does C flux relate to MAT and MAP?*

@@

MAP was a significant (p<0.05) predictor of all fluxes (Figs. 4a, S4-S5; Table S2). 
However, it explained little variation: with the exception of $R_{auto}$, MAP explained at most 25% of variation in C flux. 
For five of the nine fluxes, a second-order polynomial model provided the best fit, while first-order linear and lognormal relationships provided each provided the best fit for two flxues. 
All fluxes increased with MAP up to at least 2000 mm, above which responses were variable (Figs. 4, S4-S5).

@@

```{r echo=FALSE, out.width='100%', fig.cap = "Interactive effects of mean annual temperature and precipitation on annual forest carbon fluxes. For visualization purposes, data points are grouped into bins of 0 - 1000, 1001 - 2000, 2001 - 3000, and >3000mm mean annual precipitation, and lines of best fit models are plotted for mean annual precipitation values of 500, 1500, 2500, and 3500mm. All regressions are significant $(p<0.05)$.", fig.pos='H'}
  knitr::include_graphics("mat_map_interaction.png")
```

*How does C flux relate to other climate variables?*

Our results indicated that annual forest C fluxes were most strongly explained by temperature at the global scale, with temperature-related climate variables coming out as strong predictors of C fluxes. In addition to MAT, several of its correlates (Fig. S2) were consistently identified as strong univariate predictors of C fluxes: temperature seasonality, annual temperature range, annual frost days, PET, and length of growing season (Figs. 4, S4-S7).

All C flux variables were signficantly related to potential evapotranspiration. $ANPP_{foliage}$, $BNPP_{fine.root}$ and $R_{root}$ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (Fig. 4c, S4-5; Table S2). We found strong evidence for a saturation point or peak with PET: C fluxes tended to increase at values below 1000mm, before saturating between 1200 and 1700mm. There was also evidence that some C fluxes begin to decrease at values above 1800mm PET.

Vapour pressure deficit was a significant predictor of all C fluxes. 
$BNPP_{fine.root}$ showed a linear relationship with vapour pressure deficit ($R^2$ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (Figs. 4d, S4-5; Table S2). 
C fluxes initially increased with vapour pressure deficit, before saturating at around 0.8 kPa, after which point they began to decrease.

All fluxes, with the exception of $R_{root}$, showed a positive linear relationship with solar radiation (Figs. S4-S5, Table S2). Solar radiation explained a low proportion of variability in all C fluxes, explaining less than 20% of the variation in each flux, with the exception of $R_{auto}$ ($R^2$ = 0.26, p<0.05).

Annual wet days, cloud cover, and aridity were poor or non-significant explainers of variation in C fluxes, explaining less than 20% of the variation in each of the carbon fluxes (Figs. S4-S5; Table S2).

@@

*What is the role of seasonality in explaining C fluxes?*

Temperature seasonality was, overall, the best predictor of annual C fluxes. **(update this paragraph based on new results)**
$GPP$, $NPP$, $ANPP$, and $R_{root}$ exhibited a polynomial relationship with seasonality (all p<0.05; Figs. 4e, S6-7; Table S2). $ANPP_{foliage}$, $ANPP_{stem}$ and $R_{auto}$ decreased linearly with temperature seasonality (all p<0.05; Figs. 4e, S6-S7; Table S2). Temperature seasonality was strongly correlated with annual temperature range, which was likewise a similarly strong predictor of C fluxes (Table S2). 
C fluxes were highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower. **BBL: perhaps put this into an ecosystem context; what are these? Aseasonal subtropical places?**

In contrast, there was no significant effect of precipitation seasonality on C fluxes, and both maximum vapour pressure deficit, and water stress months were poor or non-significant predictors of variation in C fluxes (Figs. S6-S7; Table S2).

We found a significant relationship between length of growing season and C fluxes, with all fluxes showing a linear increase with length of growing season (Figs. 4e, S6-S7; Table S2). Length of growing season was a strong predictor of C fluxes, explaining 51% of variation in GPP, 39% of variation in NPP, and 34% of variation in ANPP, but it was a weaker predictor than MAT for all fluxes analysed (Table S4).

*Within the growing season, how do C fluxes vary with climate?*
